### 2022-11-09 chapters/appendix.Rmd
add chapters folder and customize scripts

---
title: "Appendix"
output: pdf_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results="hide"}
knitr::opts_chunk$set(echo = FALSE)

# load data --------------------------------------------------------------------

# Chapter 1
metro_afford_00_19 <- 
  read_rds(here("tables/metro_afford_00_19.rds"))

# set relative file path to project directory
here()

```

\appendix
\renewcommand\thefigure{A.\arabic{figure}}    
\setcounter{figure}{0}
\renewcommand\thetable{A.\arabic{table}} 
\setcounter{table}{0}

# Chapter 1 Appendix {-}

I set the appendix after the references in the main file. So, you'll just want to put all the appendix materials here. Here's an example. Appendix Table \ref{tab:metroafford} reports the median neighborhood cost burden rates for each of the 100 largest metropolitan areas between 2000 and 2019 based on their 2019 metro population size. The below text allows you to format the table and include it in the appendix. I add Latex code below to center the table and start a new page for other appendices.

\begin{center}
```{r metroafford}

# nested model object 
metro_afford_00_19 %>%

  # kable specifications
  knitr::kable(booktabs = T, 
               longtable = T,  # make long
               caption = "Median neighborhood affordability by metro, 2000 to 2019",
               linesep = "", 
               align= "lcccc", 
               col.names = c("Metro","2000","2019","2000","2019"))   %>% 
  kable_styling(latex_options = c("HOLD_position", "striped"), 
                position = "center", 
                full_width = F, 
                font_size = 9) %>%

  #  ensure equal column width of all but first column
  column_spec(c(2,3,4,5), width = "5em") %>%
  footnote(general = "Metropolitian sample is limited to the 100 largest by 2019 population. ",
           general_title = "",
           footnote_as_chunk = T)  %>%

  # add headers above each row
  add_header_above(c("", "Median rent burden" = 2, "Median mortgage burden" = 2))

```
\end{center}

\newpage