### 2022-11-09 chapters/chp1.Rmd

---
title: "Chapter 1: Insert title"
output: pdf_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results="hide"}

# remove all objects in R and clear R environment
remove(list = ls())

# chunk options ----------------------------------------------------------------
knitr::opts_chunk$set(
  fig.align = "center",        # center align figures
  fig.keep = "all",            # keep all figs
  warnings = FALSE,            # prevents warnings from appearing in code chunk
  message = FALSE,             # prevents messages from appearing in code chunk
  echo = FALSE,                # do not echo code in file
  results = "markup"           # hide results
)

# load all packages using xfun, package2 will install if not available
xfun::pkg_attach2(c("tidyverse",
                    "tidycensus",
                    "here",
                    "ggalt",
                    "ggExtra",
                    "extrafont",
                    "showtext", 
                    "kableExtra", 
                    "modelsummary",
                    "cowplot", 
                    "gt"))

# set relative file path to project directory
here()

# load data --------------------------------------------------------------------

# environment settings ---------------------------------------------------------

# set environment options
options(scipen=999)

## Loading Google fonts (https://fonts.google.com/)
font_add_google("Lato", "lato")
showtext_auto()

```

<!-- \fancyhf{}  -->
<!-- \renewcommand{\headrulewidth}{0pt}  -->
<!-- \rfoot{\thepage} -->
<!-- \pagestyle{fancy} -->
<!-- \pagenumbering{arabic} -->
\doublespacing

# Introduction {-}

Write intro.

# Section

Section 1.

## Subsection

Write something interesting.

# New section

New section. See Figure \ref{fig:dumbbellrent} that reports the median share of ... The reference in the prior sentence cross-references and auto numbers your figures. This is prob one of the most useful things about writing your dissertation in R markdown. I add latex code to singlespace around the figure below. Be sure to add a ^[footnote.]

\begin{singlespace}
```{r dumbbellrent, results = "markup", fig.cap = "Median neighborhood rent burden rate for the highest 2019 rates, 2000 to 2019", out.width = "90%"}

knitr::include_graphics(here("figures/dumbbell_rent.jpeg"))

```
\end{singlespace}

# Discussion and conclusion

Write a good conclusion.