### 2022-11-09 chapters/chp3.Rmd

---
title: "Chapter 3: Insert title"
output: pdf_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results="hide"}

# remove all objects in R and clear R environment
remove(list = ls())

# chunk options ----------------------------------------------------------------
knitr::opts_chunk$set(
  fig.align = "center",        # center align figures
  fig.keep = "all",            # keep all figs
  warnings = FALSE,            # prevents warnings from appearing in code chunk
  message = FALSE,             # prevents messages from appearing in code chunk
  echo = FALSE,                # do not echo code in file
  results = "markup"           # hide results
)

# load all packages using xfun, package2 will install if not available
xfun::pkg_attach2(c("tidyverse",
                    "tidycensus",
                    "here",
                    "ggalt",
                    "ggExtra",
                    "extrafont",
                    "showtext", 
                    "kableExtra", 
                    "modelsummary",
                    "cowplot", 
                    "gt"))

# set relative file path to project directory
here()

# load data --------------------------------------------------------------------

# functions

# environment settings ---------------------------------------------------------

# set environment options
options(scipen=999)

## Loading Google fonts (https://fonts.google.com/)
font_add_google("Lato", "lato")
showtext_auto()

```

<!-- \fancyhf{}  -->
<!-- \renewcommand{\headrulewidth}{0pt}  -->
<!-- \rfoot{\thepage} -->
<!-- \pagestyle{fancy} -->
<!-- \pagenumbering{arabic} -->
\doublespacing

# Introduction

This chapter

# Section

Here is some code to insert a figure while also centering and single spacing. Note this has a .pdf suffix instead of .png. I usually use the latter because the file size is usually smaller and easier to work with in R.

\begin{center}
\begin{singlespace}
```{r flowchart, results = "markup", fig.cap = "Declining affordability mediates the effect of gentrification on displacement", out.width = "90%"}

knitr::include_graphics(here("figures/flow-chart.pdf"))

```
\end{singlespace}
\end{center}

## Sub-section

Write some more thoughts here.

# Discussion and conclusion

Conclude it.

You can add as many chapters as you want. My dissertation had 6 chapters. Just copy and modify an .Rmd file.