### 2022-11-09 manuscript.tex
update manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{DissertateCUNY}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Georgia}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=1in,bottom=1in,right=1in,left=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{xspace}
\newcommand{\yeardegree}{2022\xspace}\newcommand{\degree}{Doctor of Philosophy\xspace}
\newcommand{\field}{Sociology\xspace}
\newcommand{\chairperson}{Chair Name\xspace}
\newcommand{\committeeone}{Prof 1\xspace}
\newcommand{\committeetwo}{Prof 2\xspace}
\newcommand{\committeethree}{Prof 3\xspace}
\newcommand{\committeefour}{\xspace}
\newcommand{\gradschoolguy}{\xspace}
\newcommand{\EO}{Executive Officer\xspace}
\newcommand{\advisor}{Advisor Name\xspace}
\newcommand{\abstract}{Write your abstract here but keep in between quotes. \xspace}
% Tables
           \usepackage{booktabs}
           \usepackage{threeparttable}
           \usepackage{array}
           \newcolumntype{x}[1]{%
           >{\centering\arraybackslash}m{#1}}%
           \usepackage{placeins}
           \usepackage{chngcntr}
           \counterwithin{figure}{chapter}
           \counterwithin{table}{chapter}
           \usepackage{lipsum}
           \usepackage[makeroom]{cancel}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdfauthor={Your Name},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Centered:\\
Title}
\author{Your Name}
\date{}

\begin{document}
\maketitle

\copyrightpage
\approvalpage
\abstractpage

\newpage
\fancyhead[L]{Acknowledgments}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}
\chapter*{ACKNOWLEDGEMENTS}
\addcontentsline{toc}{section}{Acknowledgments}

Acknowledge them.

\newpage \fancyhead[L]{Dedication} \fancyhead[R]{\thepage}
\fancyfoot[C]{}

\chapter*{DEDICATION}
\addcontentsline{toc}{section}{Dedication}

Dedicate it.

\newpage
\fancyhead[L]{Table of Contents}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}
\tableofcontents

\newpage
\fancyhead[L]{List of Tables}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}
\listoftables

\newpage
\fancyhead[L]{List of Figures}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}
\listoffigures

\newpage
\pagenumbering{arabic}

\newpage
\fancyhead[L]{Chapter 1: Running head}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}

\chapter{Chapter 1: Introduction}

\doublespacing

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Write intro.

\hypertarget{section}{%
\section{Section}\label{section}}

Section 1.

\hypertarget{subsection}{%
\subsection{Subsection}\label{subsection}}

Write something interesting.

\hypertarget{new-section}{%
\section{New section}\label{new-section}}

New section. See Figure \ref{fig:dumbbellrent} that reports the median
share of \ldots{} The reference in the prior sentence cross-references
and auto numbers your figures. This is prob one of the most useful
things about writing your dissertation in R markdown. I add latex code
to singlespace around the figure below. Be sure to add a \footnote{footnote.}

\begin{singlespace}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/dumbbell_rent}

}

\caption{Median neighborhood rent burden rate for the highest 2019 rates, 2000 to 2019}\label{fig:dumbbellrent}
\end{figure}
\end{singlespace}

\hypertarget{discussion-and-conclusion}{%
\section{Discussion and conclusion}\label{discussion-and-conclusion}}

Write a good conclusion.

\FloatBarrier

\newpage
\fancyhead[L]{Chapter 2: Running head}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}

\chapter{Chapter 2: Full title}

\doublespacing

\hypertarget{introduction-1}{%
\section{Introduction}\label{introduction-1}}

Write intro.

\hypertarget{new-section-1}{%
\section{New Section}\label{new-section-1}}

Maybe write a literature review here?

\hypertarget{sub-section}{%
\subsection{Sub-section}\label{sub-section}}

Continue writing interesting things. Add an equation. The Latex label
code allows you to cross reference the equation. I give two examples
below.

\begin{equation} \label{eq:gmi}
I = \frac{N}{W} \frac{ \Sigma_{i}\Sigma_{j}w_{ij}(x_{i} - \bar{x})(x_{j} - \bar{x}) } { \Sigma_{i}(x_{i} - \bar{x})^2 } 
\end{equation}

where \(w_{ij}\) represents the spatial weights matrix, \(N\) is the
number of spatial units denoted by \(i\) and \(j\), \(W\) is the sum of
the spatial weights, and \(x\) is the given observation (Walker 2022, pg
7.6). Essentially, this statistic summarizes the degree of spatial
autocorrelation across the spatial scope, which for this project,
represents the respective metropolitan area.

The second method combines statistical and visual evidence by
calculating Local Indicators of Spatial Autocorrelation (LISA) and
visualizing the distribution of these statistics across a metro area
using maps (Anselin 1995; Walker 2022). This affords researchers the
ability to not only visualize the spatial distribution of statistically
significant spatial clusters, but also illustrate the degree each
spatial unit contributes to overall spatial autocorrelation score
(Walker 2022). The Local Moran's I statistic takes the following form:
``a given local feature \(i\) with \(j\) neighbors:''

\begin{equation} \label{eq:lmi}
I_{i} = z_{i} \sum_{j} {w_{ij} z_{j}} 
\end{equation}

where \(i\) is a given local feature, \(j\) represents its neighbors,
and \(z_{i}\) and \(z_{j}\) denote deviations from the mean (Walker
2022).

\hypertarget{sub-sub-section.}{%
\subsubsection{Sub-sub section.}\label{sub-sub-section.}}

You can even add a sub-sub section. I find adding a ``period'' at the
end helps distinguish the section from the normal text.

\hypertarget{result-secion}{%
\section{Result secion}\label{result-secion}}

Write up your results. Here's an example of how to make a table and
cross-reference it. Table \ref{tab:descriptives} reports neighborhood
medians for all covariates included in the following multivariate
analyses by gentrification status. I include summary statistics on
neighborhoods not eligible to gentrify since these neighborhoods are
also included in the models; however, I will focus on comparing
gentrifying neighborhoods to those eligible but did not gentrify since
they are the appropriate reference group.

\begin{singlespace}
\begin{table}[H]

\caption{\label{tab:descriptives}Descriptives statistics for all model variables by gentrification status}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{15em}llll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Neighborhood gentrification status} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-4}
  & Non-gentrifying & Moderate & Intense & Not eligible\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Demographic controls}}\\
\hspace{1em}\cellcolor{gray!6}{Population} & \cellcolor{gray!6}{3725.00} & \cellcolor{gray!6}{3537.00} & \cellcolor{gray!6}{3402.50} & \cellcolor{gray!6}{3552.00}\\
\hspace{1em}Median age & 35.00 & 35.55 & 34.65 & 36.50\\
\hspace{1em}\cellcolor{gray!6}{Std. concentrated immigration} & \cellcolor{gray!6}{0.56} & \cellcolor{gray!6}{0.49} & \cellcolor{gray!6}{0.23} & \cellcolor{gray!6}{0.28}\\
\hspace{1em}Geographic mobility & 18.08 & 18.98 & 21.89 & 20.66\\
\hspace{1em}\cellcolor{gray!6}{\% non-Hispanic Black} & \cellcolor{gray!6}{20.22} & \cellcolor{gray!6}{12.09} & \cellcolor{gray!6}{5.64} & \cellcolor{gray!6}{6.00}\\
\hspace{1em}\% non-Hispanic Asian & 3.71 & 5.69 & 7.00 & 7.16\\
\hspace{1em}\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{27.26} & \cellcolor{gray!6}{21.55} & \cellcolor{gray!6}{17.11} & \cellcolor{gray!6}{15.33}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Housing market controls}}\\
\hspace{1em}Household size & 2.90 & 2.84 & 2.46 & 2.85\\
\hspace{1em}\cellcolor{gray!6}{Vacancy rate} & \cellcolor{gray!6}{7.58} & \cellcolor{gray!6}{7.68} & \cellcolor{gray!6}{8.95} & \cellcolor{gray!6}{7.79}\\
\hspace{1em}Housing market tightness & 0.85 & 0.88 & 0.90 & 0.88\\
\hspace{1em}\cellcolor{gray!6}{\% renters} & \cellcolor{gray!6}{73.06} & \cellcolor{gray!6}{67.30} & \cellcolor{gray!6}{76.60} & \cellcolor{gray!6}{63.63}\\
\hspace{1em}\% owners & 26.94 & 32.70 & 23.40 & 36.37\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Inequality controls}}\\
\hspace{1em}\cellcolor{gray!6}{Gini coefficient} & \cellcolor{gray!6}{0.44} & \cellcolor{gray!6}{0.43} & \cellcolor{gray!6}{0.46} & \cellcolor{gray!6}{0.44}\\
\hspace{1em}Std. concentrated disadvantage & 0.59 & 0.13 & -0.08 & -0.08\\
\cellcolor{gray!6}{} & \cellcolor{gray!6}{2,263} & \cellcolor{gray!6}{2,263} & \cellcolor{gray!6}{2,263} & \cellcolor{gray!6}{2,263}\\
\bottomrule
\end{tabular}
\end{table}

\end{singlespace}

\hypertarget{discussion-and-conclusion-1}{%
\section{Discussion and conclusion}\label{discussion-and-conclusion-1}}

Conclude it.

\FloatBarrier

\newpage
\fancyhead[L]{Chapter 3: Running head}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}

\chapter{Chapter 3: Full title}

\doublespacing

\hypertarget{introduction-2}{%
\section{Introduction}\label{introduction-2}}

This chapter

\hypertarget{section-1}{%
\section{Section}\label{section-1}}

Here is some code to insert a figure while also centering and single
spacing. Note this has a .pdf suffix instead of .png. I usually use the
latter because the file size is usually smaller and easier to work with
in R.

\begin{center}
\begin{singlespace}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/flow-chart}

}

\caption{Declining affordability mediates the effect of gentrification on displacement}\label{fig:flowchart}
\end{figure}
\end{singlespace}
\end{center}

\hypertarget{sub-section-1}{%
\subsection{Sub-section}\label{sub-section-1}}

Write some more thoughts here.

\hypertarget{discussion-and-conclusion-2}{%
\section{Discussion and conclusion}\label{discussion-and-conclusion-2}}

Conclude it.

You can add as many chapters as you want. My dissertation had 6
chapters. Just copy and modify an .Rmd file.

\FloatBarrier

\newpage

\fancyhead[L]{Appendices}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}
\chapter*{APPENDICES}
\addcontentsline{toc}{chapter}{APPENDICES}

\appendix
\renewcommand\thefigure{A.\arabic{figure}}    
\setcounter{figure}{0}
\renewcommand\thetable{A.\arabic{table}} 
\setcounter{table}{0}

\hypertarget{chapter-1-appendix}{%
\section*{Chapter 1 Appendix}\label{chapter-1-appendix}}
\addcontentsline{toc}{section}{Chapter 1 Appendix}

I set the appendix after the references in the main file. So, you'll
just want to put all the appendix materials here. Here's an example.
Appendix Table \ref{tab:metroafford} reports the median neighborhood
cost burden rates for each of the 100 largest metropolitan areas between
2000 and 2019 based on their 2019 metro population size. The below text
allows you to format the table and include it in the appendix. I add
Latex code below to center the table and start a new page for other
appendices.

\begin{center}
\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}[t]{l>{\centering\arraybackslash}p{5em}>{\centering\arraybackslash}p{5em}>{\centering\arraybackslash}p{5em}>{\centering\arraybackslash}p{5em}}
\caption{\label{tab:metroafford}Median neighborhood affordability by metro, 2000 to 2019}\\
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Median rent burden} & \multicolumn{2}{c}{Median mortgage burden} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Metro & 2000 & 2019 & 2000 & 2019\\
\midrule
\cellcolor{gray!6}{Akron, OH} & \cellcolor{gray!6}{33.59} & \cellcolor{gray!6}{42.59} & \cellcolor{gray!6}{24.13} & \cellcolor{gray!6}{20.32}\\
Albany-Schenectady-Troy, NY & 34.17 & 42.31 & 24.22 & 23.38\\
\cellcolor{gray!6}{Albuquerque, NM} & \cellcolor{gray!6}{40.12} & \cellcolor{gray!6}{44.39} & \cellcolor{gray!6}{30.41} & \cellcolor{gray!6}{29.20}\\
Allentown-Bethlehem-Easton, PA-NJ & 34.48 & 46.46 & 26.86 & 27.05\\
\cellcolor{gray!6}{Atlanta-Sandy Springs-Alpharetta, GA} & \cellcolor{gray!6}{35.29} & \cellcolor{gray!6}{45.23} & \cellcolor{gray!6}{24.29} & \cellcolor{gray!6}{25.26}\\
Augusta-Richmond County, GA-SC & 33.16 & 45.77 & 24.92 & 24.11\\
\cellcolor{gray!6}{Austin-Round Rock-Georgetown, TX} & \cellcolor{gray!6}{34.37} & \cellcolor{gray!6}{44.88} & \cellcolor{gray!6}{23.05} & \cellcolor{gray!6}{26.32}\\
Bakersfield, CA & 40.80 & 50.20 & 32.00 & 34.03\\
\cellcolor{gray!6}{Baltimore-Columbia-Towson, MD} & \cellcolor{gray!6}{33.89} & \cellcolor{gray!6}{45.39} & \cellcolor{gray!6}{26.67} & \cellcolor{gray!6}{26.06}\\
Baton Rouge, LA & 30.28 & 42.57 & 22.33 & 21.96\\
\cellcolor{gray!6}{Birmingham-Hoover, AL} & \cellcolor{gray!6}{29.12} & \cellcolor{gray!6}{42.30} & \cellcolor{gray!6}{23.90} & \cellcolor{gray!6}{24.42}\\
Boise City, ID & 33.92 & 41.63 & 24.40 & 25.13\\
\cellcolor{gray!6}{Boston-Cambridge-Newton, MA-NH} & \cellcolor{gray!6}{34.99} & \cellcolor{gray!6}{45.37} & \cellcolor{gray!6}{26.97} & \cellcolor{gray!6}{30.10}\\
Bridgeport-Stamford-Norwalk, CT & 36.36 & 49.57 & 32.60 & 35.36\\
\cellcolor{gray!6}{Buffalo-Cheektowaga, NY} & \cellcolor{gray!6}{37.27} & \cellcolor{gray!6}{43.70} & \cellcolor{gray!6}{25.24} & \cellcolor{gray!6}{20.77}\\
Cape Coral-Fort Myers, FL & 35.01 & 48.06 & 29.25 & 34.51\\
\cellcolor{gray!6}{Charleston-North Charleston, SC} & \cellcolor{gray!6}{33.68} & \cellcolor{gray!6}{44.06} & \cellcolor{gray!6}{26.01} & \cellcolor{gray!6}{29.33}\\
Charlotte-Concord-Gastonia, NC-SC & 29.13 & 40.44 & 24.01 & 22.92\\
\cellcolor{gray!6}{Chattanooga, TN-GA} & \cellcolor{gray!6}{30.99} & \cellcolor{gray!6}{40.91} & \cellcolor{gray!6}{23.32} & \cellcolor{gray!6}{23.04}\\
Chicago-Naperville-Elgin, IL-IN-WI & 35.00 & 45.87 & 28.68 & 30.59\\
\cellcolor{gray!6}{Cincinnati, OH-KY-IN} & \cellcolor{gray!6}{32.28} & \cellcolor{gray!6}{40.26} & \cellcolor{gray!6}{21.37} & \cellcolor{gray!6}{21.45}\\
Cleveland-Elyria, OH & 35.35 & 44.02 & 26.75 & 24.72\\
\cellcolor{gray!6}{Colorado Springs, CO} & \cellcolor{gray!6}{36.75} & \cellcolor{gray!6}{49.58} & \cellcolor{gray!6}{27.43} & \cellcolor{gray!6}{27.15}\\
Columbia, SC & 30.10 & 45.01 & 22.72 & 25.13\\
\cellcolor{gray!6}{Columbus, OH} & \cellcolor{gray!6}{31.26} & \cellcolor{gray!6}{39.43} & \cellcolor{gray!6}{23.33} & \cellcolor{gray!6}{21.77}\\
Dallas-Fort Worth-Arlington, TX & 31.85 & 44.25 & 22.13 & 25.77\\
\cellcolor{gray!6}{Dayton-Kettering, OH} & \cellcolor{gray!6}{33.75} & \cellcolor{gray!6}{40.90} & \cellcolor{gray!6}{23.00} & \cellcolor{gray!6}{20.55}\\
Deltona-Daytona Beach-Ormond Beach, FL & 38.35 & 49.40 & 32.43 & 34.04\\
\cellcolor{gray!6}{Denver-Aurora-Lakewood, CO} & \cellcolor{gray!6}{37.18} & \cellcolor{gray!6}{48.12} & \cellcolor{gray!6}{28.20} & \cellcolor{gray!6}{25.97}\\
Des Moines-West Des Moines, IA & 31.91 & 37.59 & 20.37 & 19.79\\
\cellcolor{gray!6}{Detroit-Warren-Dearborn, MI} & \cellcolor{gray!6}{33.95} & \cellcolor{gray!6}{45.49} & \cellcolor{gray!6}{22.41} & \cellcolor{gray!6}{24.40}\\
Durham-Chapel Hill, NC & 35.52 & 43.13 & 24.79 & 23.83\\
\cellcolor{gray!6}{El Paso, TX} & \cellcolor{gray!6}{38.91} & \cellcolor{gray!6}{45.48} & \cellcolor{gray!6}{29.00} & \cellcolor{gray!6}{32.98}\\
Fresno, CA & 42.57 & 50.67 & 34.11 & 32.83\\
\cellcolor{gray!6}{Grand Rapids-Kentwood, MI} & \cellcolor{gray!6}{28.04} & \cellcolor{gray!6}{41.95} & \cellcolor{gray!6}{18.25} & \cellcolor{gray!6}{19.29}\\
Greensboro-High Point, NC & 30.63 & 41.12 & 24.01 & 25.35\\
\cellcolor{gray!6}{Greenville-Anderson, SC} & \cellcolor{gray!6}{31.79} & \cellcolor{gray!6}{40.18} & \cellcolor{gray!6}{23.45} & \cellcolor{gray!6}{22.04}\\
Harrisburg-Carlisle, PA & 28.48 & 40.11 & 22.64 & 23.75\\
\cellcolor{gray!6}{Hartford-East Hartford-Middletown, CT} & \cellcolor{gray!6}{33.28} & \cellcolor{gray!6}{46.08} & \cellcolor{gray!6}{24.57} & \cellcolor{gray!6}{28.17}\\
Houston-The Woodlands-Sugar Land, TX & 31.80 & 45.29 & 22.37 & 26.38\\
\cellcolor{gray!6}{Indianapolis-Carmel-Anderson, IN} & \cellcolor{gray!6}{32.65} & \cellcolor{gray!6}{43.49} & \cellcolor{gray!6}{20.50} & \cellcolor{gray!6}{20.87}\\
Jackson, MS & 32.73 & 41.50 & 25.97 & 25.08\\
\cellcolor{gray!6}{Jacksonville, FL} & \cellcolor{gray!6}{33.33} & \cellcolor{gray!6}{46.91} & \cellcolor{gray!6}{24.93} & \cellcolor{gray!6}{27.84}\\
Kansas City, MO-KS & 30.03 & 39.95 & 19.64 & 21.08\\
\cellcolor{gray!6}{Knoxville, TN} & \cellcolor{gray!6}{31.47} & \cellcolor{gray!6}{40.61} & \cellcolor{gray!6}{24.07} & \cellcolor{gray!6}{22.49}\\
Lakeland-Winter Haven, FL & 31.13 & 45.23 & 25.25 & 28.24\\
\cellcolor{gray!6}{Las Vegas-Henderson-Paradise, NV} & \cellcolor{gray!6}{38.37} & \cellcolor{gray!6}{47.76} & \cellcolor{gray!6}{31.62} & \cellcolor{gray!6}{30.74}\\
Little Rock-North Little Rock-Conway, AR & 31.59 & 43.13 & 18.72 & 20.68\\
\cellcolor{gray!6}{Los Angeles-Long Beach-Anaheim, CA} & \cellcolor{gray!6}{43.31} & \cellcolor{gray!6}{54.80} & \cellcolor{gray!6}{41.13} & \cellcolor{gray!6}{43.01}\\
Louisville/Jefferson County, KY-IN & 29.14 & 39.70 & 20.62 & 21.71\\
\cellcolor{gray!6}{Madison, WI} & \cellcolor{gray!6}{29.05} & \cellcolor{gray!6}{38.41} & \cellcolor{gray!6}{23.73} & \cellcolor{gray!6}{23.58}\\
McAllen-Edinburg-Mission, TX & 36.36 & 45.27 & 34.97 & 32.83\\
\cellcolor{gray!6}{Memphis, TN-MS-AR} & \cellcolor{gray!6}{35.68} & \cellcolor{gray!6}{46.77} & \cellcolor{gray!6}{26.87} & \cellcolor{gray!6}{27.07}\\
Miami-Fort Lauderdale-Pompano Beach, FL & 43.92 & 58.62 & 37.30 & 41.18\\
\cellcolor{gray!6}{Milwaukee-Waukesha, WI} & \cellcolor{gray!6}{32.91} & \cellcolor{gray!6}{43.94} & \cellcolor{gray!6}{23.49} & \cellcolor{gray!6}{24.72}\\
Minneapolis-St. Paul-Bloomington, MN-WI & 33.90 & 43.23 & 20.26 & 21.83\\
\cellcolor{gray!6}{Nashville-Davidson--Murfreesboro--Franklin, TN} & \cellcolor{gray!6}{33.24} & \cellcolor{gray!6}{41.23} & \cellcolor{gray!6}{23.67} & \cellcolor{gray!6}{23.84}\\
New Haven-Milford, CT & 36.19 & 47.36 & 29.98 & 31.91\\
\cellcolor{gray!6}{New Orleans-Metairie, LA} & \cellcolor{gray!6}{38.29} & \cellcolor{gray!6}{51.03} & \cellcolor{gray!6}{27.73} & \cellcolor{gray!6}{33.15}\\
New York-Newark-Jersey City, NY-NJ-PA & 39.15 & 50.00 & 36.82 & 41.01\\
\cellcolor{gray!6}{North Port-Sarasota-Bradenton, FL} & \cellcolor{gray!6}{37.35} & \cellcolor{gray!6}{51.30} & \cellcolor{gray!6}{32.68} & \cellcolor{gray!6}{33.48}\\
Ogden-Clearfield, UT & 33.59 & 37.21 & 27.18 & 20.33\\
\cellcolor{gray!6}{Oklahoma City, OK} & \cellcolor{gray!6}{32.42} & \cellcolor{gray!6}{39.78} & \cellcolor{gray!6}{21.00} & \cellcolor{gray!6}{22.25}\\
Omaha-Council Bluffs, NE-IA & 30.43 & 40.87 & 19.32 & 21.38\\
\cellcolor{gray!6}{Orlando-Kissimmee-Sanford, FL} & \cellcolor{gray!6}{39.12} & \cellcolor{gray!6}{51.22} & \cellcolor{gray!6}{28.03} & \cellcolor{gray!6}{31.42}\\
Oxnard-Thousand Oaks-Ventura, CA & 39.92 & 54.69 & 36.50 & 39.19\\
\cellcolor{gray!6}{Palm Bay-Melbourne-Titusville, FL} & \cellcolor{gray!6}{36.93} & \cellcolor{gray!6}{46.74} & \cellcolor{gray!6}{26.84} & \cellcolor{gray!6}{31.14}\\
Philadelphia-Camden-Wilmington, PA-NJ-DE-MD & 36.56 & 46.41 & 27.80 & 28.54\\
\cellcolor{gray!6}{Phoenix-Mesa-Chandler, AZ} & \cellcolor{gray!6}{36.37} & \cellcolor{gray!6}{44.04} & \cellcolor{gray!6}{27.41} & \cellcolor{gray!6}{26.83}\\
Pittsburgh, PA & 32.69 & 38.03 & 25.00 & 20.62\\
\cellcolor{gray!6}{Portland-Vancouver-Hillsboro, OR-WA} & \cellcolor{gray!6}{37.90} & \cellcolor{gray!6}{45.42} & \cellcolor{gray!6}{30.61} & \cellcolor{gray!6}{28.47}\\
Poughkeepsie-Newburgh-Middletown, NY & 37.06 & 49.11 & 29.25 & 35.20\\
\cellcolor{gray!6}{Providence-Warwick, RI-MA} & \cellcolor{gray!6}{33.42} & \cellcolor{gray!6}{42.95} & \cellcolor{gray!6}{26.67} & \cellcolor{gray!6}{31.06}\\
Provo-Orem, UT & 35.09 & 41.53 & 30.38 & 23.28\\
\cellcolor{gray!6}{Raleigh-Cary, NC} & \cellcolor{gray!6}{32.42} & \cellcolor{gray!6}{40.91} & \cellcolor{gray!6}{22.63} & \cellcolor{gray!6}{19.85}\\
Richmond, VA & 32.61 & 45.15 & 22.73 & 26.09\\
\cellcolor{gray!6}{Riverside-San Bernardino-Ontario, CA} & \cellcolor{gray!6}{42.39} & \cellcolor{gray!6}{54.96} & \cellcolor{gray!6}{36.92} & \cellcolor{gray!6}{38.80}\\
Rochester, NY & 40.55 & 48.15 & 24.98 & 24.08\\
\cellcolor{gray!6}{Sacramento-Roseville-Folsom, CA} & \cellcolor{gray!6}{39.32} & \cellcolor{gray!6}{49.86} & \cellcolor{gray!6}{32.68} & \cellcolor{gray!6}{32.89}\\
St. Louis, MO-IL & 31.48 & 42.11 & 19.65 & 22.17\\
\cellcolor{gray!6}{Salt Lake City, UT} & \cellcolor{gray!6}{34.56} & \cellcolor{gray!6}{41.08} & \cellcolor{gray!6}{29.64} & \cellcolor{gray!6}{23.35}\\
San Antonio-New Braunfels, TX & 33.05 & 42.99 & 22.81 & 27.04\\
\cellcolor{gray!6}{San Diego-Chula Vista-Carlsbad, CA} & \cellcolor{gray!6}{41.87} & \cellcolor{gray!6}{53.25} & \cellcolor{gray!6}{38.46} & \cellcolor{gray!6}{39.19}\\
San Francisco-Oakland-Berkeley, CA & 38.30 & 43.85 & 36.97 & 35.24\\
\cellcolor{gray!6}{San Jose-Sunnyvale-Santa Clara, CA} & \cellcolor{gray!6}{38.18} & \cellcolor{gray!6}{45.65} & \cellcolor{gray!6}{34.70} & \cellcolor{gray!6}{34.46}\\
Scranton--Wilkes-Barre, PA & 30.88 & 36.50 & 27.35 & 25.54\\
\cellcolor{gray!6}{Seattle-Tacoma-Bellevue, WA} & \cellcolor{gray!6}{37.06} & \cellcolor{gray!6}{44.01} & \cellcolor{gray!6}{32.71} & \cellcolor{gray!6}{29.06}\\
Springfield, MA & 34.74 & 49.09 & 24.57 & 29.39\\
\cellcolor{gray!6}{Stockton, CA} & \cellcolor{gray!6}{42.75} & \cellcolor{gray!6}{51.73} & \cellcolor{gray!6}{35.15} & \cellcolor{gray!6}{36.00}\\
Syracuse, NY & 37.41 & 43.74 & 23.26 & 20.57\\
\cellcolor{gray!6}{Tampa-St. Petersburg-Clearwater, FL} & \cellcolor{gray!6}{37.07} & \cellcolor{gray!6}{48.65} & \cellcolor{gray!6}{27.82} & \cellcolor{gray!6}{29.86}\\
Toledo, OH & 34.18 & 39.82 & 22.22 & 22.28\\
\cellcolor{gray!6}{Tucson, AZ} & \cellcolor{gray!6}{39.41} & \cellcolor{gray!6}{45.55} & \cellcolor{gray!6}{27.28} & \cellcolor{gray!6}{28.27}\\
Tulsa, OK & 30.82 & 38.65 & 20.64 & 22.48\\
\cellcolor{gray!6}{Urban Honolulu, HI} & \cellcolor{gray!6}{38.74} & \cellcolor{gray!6}{48.72} & \cellcolor{gray!6}{40.41} & \cellcolor{gray!6}{37.44}\\
Virginia Beach-Norfolk-Newport News, VA-NC & 37.54 & 47.84 & 30.23 & 31.09\\
\cellcolor{gray!6}{Washington-Arlington-Alexandria, DC-VA-MD-WV} & \cellcolor{gray!6}{31.74} & \cellcolor{gray!6}{43.94} & \cellcolor{gray!6}{25.12} & \cellcolor{gray!6}{26.57}\\
Wichita, KS & 30.75 & 41.05 & 18.36 & 21.47\\
\cellcolor{gray!6}{Winston-Salem, NC} & \cellcolor{gray!6}{28.36} & \cellcolor{gray!6}{42.16} & \cellcolor{gray!6}{23.23} & \cellcolor{gray!6}{23.25}\\
Worcester, MA-CT & 30.95 & 41.57 & 23.94 & 27.61\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}Metropolitian sample is limited to the 100 largest by 2019 population. }\\
\end{longtable}
\endgroup{}
\end{center}

\newpage

\FloatBarrier

\newpage

\fancyhead[L]{References}
\fancyhead[R]{\thepage}
\fancyfoot[C]{}

\chapter*{REFERENCES}
\singlespacing

\setlength{\parskip}{6pt}

\noindent

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Anselin1995}{}}%
Anselin, Luc. 1995. {``Local {Indicators} of {Spatial
Association}\textemdash{{LISA}}.''} \emph{Geographical Analysis}
27(2):93--115. doi:
\href{https://doi.org/10.1111/j.1538-4632.1995.tb00338.x}{10.1111/j.1538-4632.1995.tb00338.x}.

\leavevmode\vadjust pre{\hypertarget{ref-Walker2022}{}}%
Walker, Kyle. 2022. \emph{Analyzing {US Census Data}: {Methods}, {Maps},
and {Models} in {R}}. {CRC Press Taylor \& Francis Group}.

\end{CSLReferences}

\clearpage

\end{document}